master thesis

### 2018-04-01 Paper - Latex/Paper - Skizze - v4.tex
Thesis - Appendix

\begin{document}
  \begin{titlepage}
    \pagestyle{empty}
  	\begin{center}
      {\Large Freie Universität Berlin} \\
    	\begin{Huge}
      	Fachbereich Mathematik und Informatik \\
      	\vspace{3mm}
    	\end{Huge}
    	\vspace{2cm}
    	\begin{Large}
        \textbf{An analysis of maximum parsimony algorithms to predict parasitism in Eukaryota} \\
        \vspace{3mm}
        using a large multifurcated phylogenetic synthesis tree \\
    	\end{Large}
      \vspace{3cm}
      \textbf{Submitted on:} \\
      3 April 2018 \\
    	\vspace{2cm}
    	Lydia Buntrock \\
      E-Mail: info@irallia.de \\
     	\vspace{3cm}
      \textbf{Supervisor:} \\
      Prof. Dr. Bernhard Y. Renard \\
      \& \\
      Prof. Dr. rer. nat. Emanuel Heitlinger \\      
  	\end{center}
    \clearpage
    \pagenumbering{Roman}
  \end{titlepage}

@@

  One of the major disadvantages of parsimony methods is that, unlike likelihood approaches, they 
    can not take divergence times (branch length) into account. Since the OTL does not include 
    development times of species, this can not be considered here. \\
  Another problem pointed out by Royer-Carenzi is that parsimony approaches are either based on 
    predefined parameters (generalized parsimony) or on strong and often controversial assumptions, 
    like irreversibility of transitions for dollo parsimony. This problem is unimportant to the 
    present task because in the analysis of the entire Eukaryota tree only generalized models are 
    meaningful. \\
  % Comparing the methods, we find that parsimony are methods in which you can not include information 
  %   such as branch length (development time of the species). This is also their main disadvantage 
  %   compared to the more complex likelihood based models. Since there are no branch lengths or other 
  %   information available for the present relationship tree, and we can thus not take advantage of 
  %   the other models. \\
  The comparison of methods gives us the maximum parsimony method as the simplest and fastest method 
    that meets all our requirements.
  % Parsimony-based methods are used in this work, since they are fast and fully sufficient for the 
  %   presented use case here, following the principle of the simpler model first. \\

@@

  There are many different ways to define parasitism. Since we use GloBI to classify species, we use 
    their definition of parasitism. Again, in GloBi, Ontobee definitions are used \cite{Xiang2011}. \\
  The interaction \textit{has parasite} is defined as: "An interaction relationship between two 
    organisms living together in more or less intimate association in a relationship in which 
    association is disadvantageous or destructive to one of the organisms."\footnote{
      \hyperlink{http://www.ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO_0002445}
      {ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO\_000244}; Last checked: 22.03.2018.
    }. This definition includes: ecto- and endoparasites, parasitoids, kleptoparasites and pathogenes. \\

@@

  For this reconstruction, we first analyzed our data. This is the tree of OTL \cite{Hinchliff2015}
    and the data for the leaf node states (free-living or parasite states) from GloBI 
    \cite{Poelen2014} (section \ref{sec:results - missing information}), where we have determined 
    that the entire Eukaryota tree is 89.45\% multifurcated (missing internal nodes) and that we do 
    not have any state information for 97.34\% nodes. \\
  In addition, we tested different models for these two paramters and found out that multifurcation 
    can best be modeled with the taxa 'order' multiplied by the mean height of the knots. Also, the 
    missing states could be modeled with the taxa 'order' and in this case multiplied by the depth 
    of the nodes.

@@

  As a result, we used the Sankoff algorithm \cite{Sankoff1975} implemented by Louca et al. 
    \cite{Louca2017}, which performed best, for the actual reconstruction on the real data. \\
  We evaluated these on the one hand by analysing some statisitcs (subsection 
    \ref{subsec:results - statistics on predicted states}) and on the other hand by performing a 
    leave-100-out cross-validation. There we predicted approximately 98.17\% of the omitted states 
    correct (subsection \ref{sec:results - cross-validation}). \\

@@

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models are created with the R function \textit{glm()} and compared with the \textit{BIC()} 
            function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC multifurcation} 
      \end{table}
      Within every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (e.g. 
        family) are computationally too expensive to calculate. \\
      The model \textit{multifurc $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that e.g. \textit{multifurc $\sim$ family * mean.height} is better. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for missing state information}
      Next to the problem of the multifurcation of the tree is the little interaction data thet we have 
        for the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The Eukaryota synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 free-living species and 47,332 parasite species 
        for the whole tree of life. From the Eukaryota we could determine 25,962 and 34,860 species as 
        parasites and free-living. With 2,293,463 leaf nodes we still have about 97.34\% unknown leaf 
        nodes.

      We also compared different models in terms of their BICs (table: \ref{table:BIC unknown information}). \\ 
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC unknown information} 
      \end{table}

@@

      Similar to the Eukaryota it is with the Metazoa subtree, which accounts for most of the Eukayota 
        (179,944 internal and 1,491,012 leaf nodes). However, here the root node is no longer calculated 
        as parasitic but still as indefinable. \\
      Strongly parasitic subtrees are for example the Apicomplexa, the Nematoda and the Platyhelminthes. 
        For all three, the prediction is at a similar proportion between input values of the states and 
        predicted states, e.g. In the Apicomplexa with 99.61\% known parasitic states, 99.95\% parasites 
        were predicted. The same applies to the chordata, only that these mainly consist of free-living 
        species.

@@

          \hspace*{-0.5cm}\begin{tabular}{|>{\scriptsize}c|>{\scriptsize}l|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r| }

@@

    For one thing, there were a few subtrees of higher ranks whose information share was much higher. 
      In the Chordata there is only 66.49\% multifurcation and in the Platyhelminthes only 68.73\% 
      missing state information. \\
    Furthermore, we know that OTL is a synthesis of 819 phylogenetic trees. This suggests that just 
      these parts of the tree are very well resolved and the remainder consists only of the rough 
      taxonomy. \\
    \todotext{
    For a review of this underlying taxonomy, we have compiled a list of all taxa and drawn them in 
      relation to their distribution (see appendix section \ref{sec:otl analysis}).
    The number of nodes in the tree corresponding to a taxonomic rank is tabulated. \\
    For the classical taxonomic ranks phylum, class, order, family, genus xyz is found (appendix 
      \ref{sec:otl analysis}). \\
    A closer look is taken at some of the features of the synthesis tree on the one hand the 
      distribution of the taxa and on the other the distribution of the nodes on the taxa. Since 
      this is \colorbox{red}{not directly relevant} for this study, this section has been moved to 
      the appendix \ref{sec:otl analysis}.}

@@

    %---------------------------------------------------------------------------------------------------
    \subsection{Distribution of Taxa}
    In the tree we can distinguish 28 different Taxa with the OTL taxonomic tree.
    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\textwidth]{Figures/TaxaTable2.JPG}
      \caption{Distribution of Nodes in Rank-Cathegories}
      \label{fig:taxaTable2}
    \end{figure}

    \begin{table}[h!]
      \begin{center}
        % \begin{tabular}{ |>{\rowmac}l|>{\rowmac}r|<{\clearrow} }
        \begin{tabular}{ |l|r||r|r|r| }
          \hline
          \bfseries Taxa & \bfseries Number of Nodes & \bfseries Internal Nodes & \bfseries Leaf Nodes & \bfseries Terminal Leaf Nodes \\
          \hline \hline
          \setrow{\bfseries}domain & 1 & 1 & &  \\ \hline
          \setrow{\bfseries}kingdom & 3 & 3 & &  \\
          subkingdom & 2 & 2 & & \\
          infrakingdom & 5 & 5 & & \\
          superphylum & 2 & 2 & & \\ \hline
          \setrow{\bfseries}phylum & 57 & 54 & 3 & \\
          subphylum & 22 & 20 & 2 & \\
          infraphylum & 2 & 2 & & \\
          superclass & 8 & 8 & & \\ \hline
          \setrow{\bfseries}class & 209 & 175 & 34 & \\
          subclass & 120 & 106 & 14 & \\
          infraclass & 14 & 11 & 3 & \\
          superorder & 40 & 37 & 3 & \\ \hline
          \setrow{\bfseries}order & 1014 & 772 & 242 & \\
          suborder & 285 & 222 & 63 & \\
          infraorder & 95 & 78 & 17 & \\
          parvorder & 14 & 12 & 2 & \\
          superfamily & 829 & 626 & 203 & \\ \hline
          \setrow{\bfseries}family & 8449 & 6143 & 2306 & \\
          subfamily & 3090 & 2250 & 840 & \\
          supertribe & 1 & 0 & 1 & \\
          tribe & 1764 & 1285 & 479 & \\
          subtribe & 435 & 359 & 77 & \\ \hline
          \setrow{\bfseries}genus & 164656 & 105452 & 59204 & \\
          subgenus & 1266 & 869 & 397 & \\
          section & 3 & 2 & 1 & \\
          subsection & 1 & 1 & 0 & \\
          species group & 365 & 288 & 77 & \\
          species subgroup & 123 & 100 & 23 & \\ \hline
          \setrow{\bfseries}species & 2247251 & 7423 & 2239828 & 2228993 \\
          subspecies & 22437 & 75 & 22362 & 22239 \\
          variety & 1755 & 7 & 1748 & 1726 \\
          varietas & 5970 & 19 & 5951 & 5909 \\
          forma & 1181 & & 1181 & 1181 \\
          \hline \hline
          no rank & 954 & 719 & 235 & 7 \\
          no rank - terminal & 37452 & & 37452 & 37452 \\
          (no entry) & 40099 & 40099 & & \\
          \hline  
        \end{tabular}
        \caption{\todo{...}}
        \label{table:...} 
      \end{center}  
    \end{table}
    extended leaf nodes (real leaf nodes)

  %---------------------------------------------------------------------------------------------------
  \subsection{Distribution of data in the taxa}
    Mithilfe des taxonomischen Baums von OTL haben wir die Knoten ihren Kingdoms, Phyla und Classes 
      zugeteilt (see table \ref{table:...}). \\

    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l|r|l|r|r| }
          \hline
          \bfseries Kingdom (3) & \bfseries Number of Nodes & \bfseries Phylum (25) & \bfseries Number of Nodes & \bfseries max max height \\
          \hline \hline
          Metazoa & 1 465 207 & Arthropoda & 1 170 539 & 54 \\
          & & Chordata & 106 650 & 74 \\
          & & Mollusca & 80 022 & 22 \\
          & & Platyhelminthes & 27 141 & 9 \\
          & & Nematoda & 24 564 & 23 \\
          & & Cnidaria & 14 878 & 36 \\
          & & Porifera & 11 737 & 26 \\
          & & Echinodermata & 10 654 & 14 \\
          & & Bryozoa & 8 631 & 11 \\
          & & Rotifera & 3 093 & 7 \\
          & & Nemertea & 1 793 & 7 \\
          & & Tardigrada & 1 654 & 7 \\
          & & Acanthocephala & 1 596 & 6 \\
          & & Brachiopoda & 1 055 & 9 \\
          & & Nematomorpha & 633 & 7 \\
          & & Chaetognatha & 360 & 7 \\
          & & Hemichordata & 196 & 5 \\ 
          & & Cycliophora & 11 & 5 \\ 
          % 1465207-(1170539+106650+80022+27141+24564+14878+11737+10654+8631+3093+1793+1654+1596+1055+633+360+196+11) = 0
          \hline
          Fungi & 254 871 & Ascomycota & 157 045 & 19 \\ 
          & & Basidiomycota & 92 931 & 18 \\
          & & Microsporidia & 1 949 & 6 \\ 
          & & Glomeromycota & 1 490 & 6 \\
          & & Chytridiomycota & 1 456 & 6 \\
          % 254871-157045-92931-1949-1490-1456 = 0
          \hline
          Chloroplastida & 121 239 & Streptophyta & 120 731 & 49 \\
          & & Chlorophyta & 508 & 6 \\
          % 121239-120731-508 = 0
          \hline  
        \end{tabular}
        \caption{\todo{...}}
        \label{table:...} 
      \end{center}  
    \end{table}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Parasites in Chordata} \label{sec:parasites in chordata}

    \begin{figure}
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.85\textwidth]{Figures/ChordataParasites.png}
      \caption{Parasites of Phylum: Chordata. \\
        All parasite data of the chordata are mapped into a rough taxonomy (phylum, class, order, 
          family) in order to understand their distribution and affiliation. \\
        The internal nodes are the wanted taxa from OTL (with the addition of \# input parasites to 
          $\rightarrow$ \# predicted parasites). \\
        The leaf nodes are the input parasites (green) and the predicted parasites (white $\rightarrow$ 
          green).}
      \label{fig:ChordataParasites}
    \end{figure}

    We mapped the few parasitic species in a rough taxonomy (see Figure \ref{fig:ChordataParasites}): \\
    Known parasitic birds belong to the order Sauria. Here we know from Rothschild here that are 
      breeding parasites, like the cuckoo and clepto-parasites as the skuas \cite{Rothschild1957}. We
      got 6 input parasites from GloBI and there are no predictions: A woodpecker - 
      \textit{Sphyrapicus varius} and a duck - \textit{Aix sponsa}, a cow bird - \textit{Molothrus ate} 
      known as broodparasite and some others. \\
    An example of the amplification of mistakes here are the carp. There is a paper from which GloBI 
      concludes: Grass carp (\textit{Ctenopharyngodon idella)} has Pathogen common carp 
      (\textit{Cyprinus carpio})\footnote{
        \hyperlink{https://www.globalbioticinteractions.org/?interactionType=hasParasite&targetTaxon=Cyprinus\%20carpio}
        {https://www.globalbioticinteractions.org/?interactionType=hasParasite\&targetTaxon=Cyprinus\%20carpio}
      }. Since there is hardly any information about free-living, it follows that all siblings are 
      also predicted to be parasitic.

>>>>

\begin{document}
  \begin{titlepage}
    \pagestyle{empty}
  	\begin{center}
      {\Large Freie Universität Berlin} \\
    	\begin{Huge}
      	Fachbereich Mathematik und Informatik \\
      	\vspace{3mm}
    	\end{Huge}
    	\vspace{2cm}
    	\begin{Large}
        \textbf{An analysis of maximum parsimony algorithms to predict parasitism in Eukaryota} \\
        \vspace{3mm}
        using a large multifurcated phylogenetic synthesis tree \\
    	\end{Large}
      \vspace{3cm}
      \textbf{Submitted on:} \\
      3 April 2018 \\
    	\vspace{2cm}
    	Lydia Buntrock \\
      E-Mail: info@irallia.de \\
     	\vspace{3cm}
      \textbf{Supervisors:} \\
      Prof. Dr. Bernhard Y. Renard \\
      \& \\
      Prof. Dr. rer. nat. Emanuel Heitlinger \\      
  	\end{center}
    \clearpage
    \pagenumbering{Roman}
  \end{titlepage}

@@

  One of the major disadvantages of parsimony methods is that, unlike likelihood approaches, they 
    can not take divergence times (branch length) into account. Since the OTL does not include 
    development times of species, this can not be considered here. \\
  Another problem pointed out by Royer-Carenzi is that parsimony approaches are either based on 
    predefined parameters (generalized parsimony) or on strong and often controversial assumptions, 
    like irreversibility of transitions for Dollo parsimony. This problem is unimportant to the 
    present task because in the analysis of the entire Eukaryota tree only generalized models are 
    meaningful. \\
  % Comparing the methods, we find that parsimony are methods in which you can not include information 
  %   such as branch length (development time of the species). This is also their main disadvantage 
  %   compared to the more complex likelihood based models. Since there are no branch lengths or other 
  %   information available for the present relationship tree, and we can thus not take advantage of 
  %   the other models. \\
  The comparison of methods gives us the maximum parsimony method as the simplest and fastest method 
    that meets all our requirements.
  % Parsimony-based methods are used in this work, since they are fast and fully sufficient for the 
  %   presented use case here, following the principle of the simpler model first. \\

@@

  There are many different ways to define parasitism. Since we use GloBI to classify species, we use 
    their definition of parasitism. Again, in GloBi, Ontobee definitions are used \cite{Xiang2011}. \\
  The interaction \textit{has parasite} is defined as: "An interaction relationship between two 
    organisms living together in more or less intimate association in a relationship in which 
    association is disadvantageous or destructive to one of the organisms."\footnote{
      \hyperlink{http://www.ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO_0002445}
      {ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO\_000244}; Last checked: 22.03.2018.
    }. This definition includes: ecto- and endoparasites, parasitoids, kleptoparasites and pathogens. \\

@@

  For this reconstruction, we first analyzed our data. This is the tree of OTL \cite{Hinchliff2015}
    and the data for the leaf node states (free-living or parasite states) from GloBI 
    \cite{Poelen2014} (section \ref{sec:results - missing information}), where we have determined 
    that the entire Eukaryota tree is 89.45\% multifurcated (missing internal nodes) and that we do 
    not have any state information for 97.34\% nodes. \\
  In addition, we tested different models for these two parameters and found out that multifurcation 
    can best be modeled with the taxa 'order' multiplied by the mean height of the knots. Also, the 
    missing states could be modeled with the taxa 'order' and in this case multiplied by the depth 
    of the nodes.

@@

  As a result, we used the Sankoff algorithm \cite{Sankoff1975} implemented by Louca et al. 
    \cite{Louca2017}, which performed best, for the actual reconstruction on the real data. \\
  We evaluated these on the one hand by analysing some statistics (subsection 
    \ref{subsec:results - statistics on predicted states}) and on the other hand by performing a 
    leave-100-out cross-validation. There we predicted approximately 98.17\% of the omitted states 
    correct (subsection \ref{sec:results - cross-validation}). \\

@@

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurcation $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurcation $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurcation $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurcation $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurcation $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurcation $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurcation $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurcation $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurcation $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurcation $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models are created with the R function \textit{glm()} and compared with the \textit{BIC()} 
            function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC multifurcation} 
      \end{table}
      Within every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (e.g. 
        family) are computationally too expensive to calculate. \\
      The model \textit{multifurcation $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that e.g. \textit{multifurcation $\sim$ family * mean.height} is better. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for missing state information}
      Next to the problem of the multifurcation of the tree is the little interaction data that we have 
        for the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The Eukaryota synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 free-living species and 47,332 parasite species 
        for the whole tree of life. From the Eukaryota we could determine 25,962 and 34,860 species as 
        parasites and free-living. With 2,293,463 leaf nodes we still have about 97.34\% unknown leaf 
        nodes.

      We also compared different models in terms of their BICs (table: \ref{table:BIC unknown information}). \\ 
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurcation $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurcation $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurcation $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC unknown information} 
      \end{table}

@@

      Similar to the Eukaryota it is with the Metazoa subtree, which accounts for most of the Eukaryota 
        (179,944 internal and 1,491,012 leaf nodes). However, here the root node is no longer calculated 
        as parasitic but still as indefinable. \\
      Strongly parasitic subtrees are for example the Apicomplexa, the Nematoda and the Platyhelminthes. 
        For all three, the prediction is at a similar proportion between input values of the states and 
        predicted states, e.g. In the Apicomplexa with 99.61\% known parasitic states, 99.95\% parasites 
        were predicted. The same applies to the chordata, only that these mainly consist of free-living 
        species.

@@

          \hspace*{-0.5cm}\begin{tabular}
              {|>{\scriptsize}c|>{\scriptsize}l|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r||
                >{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r| }

@@

    For one thing, there were a few subtrees of higher ranks whose information share was much higher. 
      In the Chordata there is only 66.49\% multifurcation and in the Platyhelminthes only 68.73\% 
      missing state information. \\
    Furthermore, we know that OTL is a synthesis of 819 phylogenetic trees. This suggests that just 
      these parts of the tree are very well resolved and the remainder consists only of the rough 
      taxonomy. For a review of this underlying taxonomy, we have compiled a list of all taxa and 
      plottet them in relation to their distribution (see appendix section \ref{sec:otl analysis}). 
      There the number of nodes in the tree corresponding to a taxonomic rank is tabulated.

@@

    %---------------------------------------------------------------------------------------------------
    \subsection{Distribution of Taxa}
      In the tree we can distinguish 28 different Taxa in the OTL taxonomic tree. We have all Taxa 
        sorted by rank into a table and counted their occurrences as internal nodes and leaf nodes
        (see table \ref{table:taxa}) and plotted them as a diagram (see figure \ref{fig:taxa}).
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\textwidth]{Figures/TaxaTable2.JPG}
      \caption{Distribution of Nodes in Rank-Cathegories}
      \label{fig:taxa}
    \end{figure}

    \begin{table}[h!]
      \begin{center}
        % \begin{tabular}{ |>{\rowmac}l|>{\rowmac}r|<{\clearrow} }
        \begin{tabular}{ |l|r||r|r| }
          \hline
          \bfseries Taxa & \bfseries Number of Nodes & \bfseries Internal Nodes & \bfseries Leaf Nodes \\
          \hline \hline
          \setrow{\bfseries}domain & 1        & 1 & - \\ \hline
          \setrow{\bfseries}kingdom & 3       & 3 & - \\
          subkingdom & 2                      & 2 & - \\
          infrakingdom & 5                    & 5 & - \\
          superphylum & 2                     & 2 & - \\ \hline
          \setrow{\bfseries}phylum & 57       & 57 & - \\
          subphylum & 22                      & 22 & - \\
          infraphylum & 2                     & 2 & - \\
          superclass & 8                      & 8 & - \\ \hline
          \setrow{\bfseries}class & 209       & 209 & - \\
          subclass & 120                      & 120 & - \\
          infraclass & 14                     & 14 & - \\
          superorder & 40                     & 40 & - \\ \hline
          \setrow{\bfseries}order & 1013      & 1013 & - \\
          suborder & 285                      & 285 & - \\
          infraorder & 95                     & 95 & - \\
          parvorder & 14                      & 14 & - \\
          superfamily & 829                   & 829 & - \\ \hline
          \setrow{\bfseries}family & 8442     & 8442 & - \\
          subfamily & 3090                    & 3090 & - \\
          supertribe & 1                      & 1 & - \\
          tribe & 1764                        & 1764 & - \\
          subtribe & 435                      & 435 & - \\ \hline
          \setrow{\bfseries}genus & 164592    & 164592 & - \\
          subgenus & 1266                     & 1266 & - \\
          section & 3                         & 3 & - \\
          subsection & 1                      & 1 & - \\
          species group & 308                 & 308 & - \\
          species subgroup & 108              & 108 & - \\ \hline
          \setrow{\bfseries}species & 2243241 & 18197 & 2225044 \\
          subspecies & 22378                  & 196 & 22182 \\
          variety & 1755                      & 29 & 1726 \\
          varietas & 5935                     & 61 & 5874 \\
          forma & 1179                        & & 1179 \\
          \hline \hline
          no rank & 951                       & 944 & 7 \\
          no rank - terminal & 37451          & - & 37451 \\
          (no entry) & 39816                  & 39816 & - \\
          \hline  
        \end{tabular}
        \caption{Distribution of Nodes in Rank-Cathegories}
        \label{table:taxa} 
      \end{center}  
    \end{table}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Parasites in Chordata} \label{sec:parasites in chordata}
    We mapped the few parasitic species in a rough taxonomy (see Figure \ref{fig:ChordataParasites}): \\
    Known parasitic birds belong to the order Sauria. Here we know from Rothschild here that are 
      breeding parasites, like the cuckoo and clepto-parasites as the skuas \cite{Rothschild1957}. We
      got 6 input parasites from GloBI and there are no predictions: A woodpecker - 
      \textit{Sphyrapicus varius} and a duck - \textit{Aix sponsa}, a cow bird - \textit{Molothrus ate} 
      known as broodparasite and some others. \\
    An example of the amplification of mistakes here are the carp. There is a paper from which GloBI 
      concludes: Grass carp (\textit{Ctenopharyngodon idella)} has Pathogen common carp 
      (\textit{Cyprinus carpio})\footnote{
        \hyperlink{https://www.globalbioticinteractions.org/?interactionType=hasParasite&targetTaxon=Cyprinus\%20carpio}
        {https://www.globalbioticinteractions.org/?interactionType=hasParasite\&targetTaxon=Cyprinus\%20carpio}
      }. Since there is hardly any information about free-living, it follows that all siblings are 
      also predicted to be parasitic.

    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.85\textwidth]{Figures/ChordataParasites.png}
      \caption{Parasites of Phylum: Chordata. \\
        All parasite data of the chordata are mapped into a rough taxonomy (phylum, class, order, 
          family) in order to understand their distribution and affiliation. \\
        The internal nodes are the wanted taxa from OTL (with the addition of \# input parasites to 
          $\rightarrow$ \# predicted parasites). \\
        The leaf nodes are the input parasites (green) and the predicted parasites (white $\rightarrow$ 
          green).}
      \label{fig:ChordataParasites}
    \end{figure}
